A Non-local Bayesian Video Denoising Method

### 2015-10-19 doc/201507_tip/paper.tex
[doc] Draft 1 of paper, sent to JMM for feedback.

The rank also plays a role in the estimation of the covariance matrix.
As explained in \S \ref{sec:complexity}, alleviating the computational cost is
one of the reasons for using a low rank approximation of the empirical
covariance matrix. But there are also good reasons regarding the denoising
performance. In situations in which the number of samples 
is comparable to their dimension, the eigen-decomposition of the empirical
covariance matrix $\widehat C$ can have considerable errors with respect to
that of the population covariance matrix $C$. 
%
The case of Gaussian models with covariance matrices of the form $C = A + \sigma^2I$,
where $A$ is does not have full rank, has been studied in the statistics and physics
literature. The authors in \cite{johnstone2009,paul2007,nadler2008} study the
limit when $n,p\to\infty$ such that $p/n\to\gamma < \infty$.
These works show that, for a non-zero ratio $\gamma$, the empirical eigenvectors 
and eigenvalues are inconsistent estimators of the population ones.
Furthermore, the smaller eigenvalues and their eigenvectors will have a higher
error and eventually, if the signal-to-noise ratio drops bellow $\sqrt{\gamma}$, 
the eigenvector is confused with the noise and the resulting estimate will be,
in expectation, orthogonal to the actual value.
Finite sample bounds for the eigenvalues and eigenvectors have been also studied 
by \cite{nadler2008}, in the case in which $A$ has rank 1.
\pcomment{What can we say about this bounds.}
These larger errors between the sample eigenvectors and their population
counterparts manifest themselves as noise, which then affects the MAP estimate.
This suggest that better denoising should be obtained by discarding the
eigenvectors associated with lower eigenvalues.

@@

For each stage of the algorithm we need to specify 6 parameters: the spatial
and temporal components of the patch size, $s_x$ and $s_t$, the spatial and
temporal components of the search region, $w_x$ and $w_t$, the number of similar
patches $n$ and the rank of the a priori covariance matrix $r$.
When needed, we will add a subscript indicating the stage, \eg
$n_{1}$ and $n_{2}$.

@@

Setting the size of the search region implies a
trade-off between denoising performance and running time: A larger search
region takes longer to traverse but yields better results, since more
similar patches can be found. We fix the search
region for both stages of the algorithm at 
\begin{equation}
w_x = 37 \quad \text{ and } \quad w_t = 5.
\label{eq:search_region_parameters}
\end{equation}
With these values, the search region captures the $5$ frame trajectory of the
target patch as long as it moves slower than $37/5 = 7.4$
pixels/frame.

@@

We will set the spatial patch size at \[s_x = 5\] for both stages. Larger
patches (up to $s_x = 9$) lead to an mild improvement of the results (on
average). This improvement comes at a considerable increase in the
computation time, since our current implementation is $\mathcal O(s_x^4)$. Thus,
have found that a good compromise between computation time and quality is given
by $s_x = 5$.

@@

Due the coupling of the first and second stages, determining the optimal values
for the ranks $r_1, r_2$ and the number of patches $n_1,n_2$ requires a four
dimensional parameter search.
%
Instead, we fix the first stage parameters at
\[r_1 = s_{x,1}^2s_{t,1}\quad\quad n_{1} = 400,\]
which we found to give good results. In doing so, we are being conservative and
use the full rank covariance matrix to eliminate the least signal components.
%
We then run a parameter search to determine $r_2$ and $n_{2}$.
The results is the ones shown in Figure \ref{fig:rank1-nsim1-fpsnr}. From here
establish the following values as a function of $\sigma$:
%\[r_2(\sigma) = 40 - \frac13(\sigma - 10)\quad\quad n_{2} = 150.\]
\[r_2(\sigma) = 35 - \frac12(\sigma - 10)\quad\quad n_{2} = 100.\]
Note that a higher value for $n_{2}$ would give slightly worse results,
at a lower computational cost.

@@

		\vspace{1cm}
		\pcomment{TODO: Figures for grayscale sequences}
		\vspace{1cm}

	\end{center}
	\caption{Basic estimate PSNR, averaged over four sequences, when varying the
		second stage parameters $r_1$ and $n_1$. From left to right, results for
		$\sigma = 10, 20, 30, 40$. In the top row: results for RGB sequences, bottom
		row, results for grayscale sequences.} 
	\label{fig:rank1-nsim1-bpsnr}
\end{figure*}

@@

		\vspace{1cm}
		\pcomment{TODO: Figures for grayscale sequences}
		\vspace{1cm}

	\end{center}
	\caption{PSNRs of final vs. basic estimate. Averages over four sequences, when varying the
		second stage parameters $r_1$ and $n_1$. From left to right, results for
		$\sigma = 10, 20, 30, 40$. Top: results for RGB sequences; bottom:
		results for grayscale sequences. The width of the line is proportional to
		the rank. \pcomment{These plots contradict the argument in favour of using
		full rank in the first stage (supposedly introducing a higher variance, but a
		smaller bias). It turns out, that (1) the final PSNR is pretty much
		correlated with the basic PSNR, and (2) the optimal rank for the final
		estimate is lower than that of the basic estimate. I do not have an explanation for this.}} 
	\label{fig:rank1-nsim1-bpsnr-fpsnr}
\end{figure*}

@@

		\vspace{1cm}
		\pcomment{TODO: Figures for grayscale sequences}
		\vspace{1cm}

	\end{center}
	\caption{Final estimate PSNR, averaged over four sequences, when varying the
		second stage parameters $r_1$ and $n_1$. From left to right, results for
		$\sigma = 10, 20, 30, 40$. In the top row: results for RGB sequences, bottom
		row, results for grayscale sequences.} 
	\label{fig:rank1-nsim1-fpsnr}
\end{figure*}

@@

		\vspace{1cm}
		\pcomment{TODO: Figures for grayscale sequences}
		\vspace{1cm}

@@

		{%\small

@@

	\caption{PSNRs obtained for the four classic color test sequences used in
	\cite{Maggioni2012}. See text for details.}

@@

	\caption{Quantitative denoising results for some classic grayscale test sequences. See text for details.}

@@

In this section we present results obtained with the proposed
method on some classical test sequences with Gaussian noise added.
We consider separatedly the cases of grayscale and color videos.

@@

	Sintel's database. See text for details.}

>>>>

The rank also plays a role in the estimation of the covariance matrix.
As explained in \S \ref{sse:complexity}, alleviating the computational cost is
one of the reasons for using a low rank approximation of the empirical
covariance matrix. But there are also good reasons regarding the denoising
performance. In situations in which the number of samples 
is comparable to their dimension, the eigen-decomposition of the empirical
covariance matrix $\widehat C$ can have considerable errors with respect to
that of the population covariance matrix $C$. 
%
The case of Gaussian models with covariance matrices of the form $C = A + \sigma^2I$,
where $A$ is does not have full rank, has been studied in the statistics and physics
literature. The authors in \cite{johnstone2009,paul2007,nadler2008} study the
limit when $n,p\to\infty$ such that $p/n\to\gamma < \infty$.
These works show that, for a non-zero ratio $\gamma$, the empirical eigenvectors 
and eigenvalues are inconsistent estimators of the population ones.
Furthermore, the smaller eigenvalues and their eigenvectors will have a higher
error and eventually, if the signal-to-noise ratio drops bellow $\sqrt{\gamma}$, 
the eigenvector is confused with the noise and the resulting estimate will be,
in expectation, orthogonal to the actual value.
Finite sample bounds for the eigenvalues and eigenvectors have been also studied 
by \cite{nadler2008}, in the case in which $A$ has rank 1.
\pcomment{What can we say about this bounds.}
These larger errors between the sample eigenvectors and their population
counterparts manifest themselves as noise, which then affects the MAP estimate.
This suggest that better denoising should be obtained by discarding the
eigenvectors associated with lower eigenvalues.

@@

For each stage of the algorithm we need to specify 6 parameters: the spatial
and temporal components of the patch size, $s_x$ and $s_t$, the spatial and
temporal components of the search region, $w_x$ and $w_t$, the number of similar
patches $n$ and the rank of the a priori covariance matrix $r$.
When needed, we will add a subscript indicating the stage, \eg
$n_{1}$ and $n_{2}$.
We will consider different parameter settings for color and grayscale videos.

@@

Setting the size of the search region implies a trade-off between denoising
performance and running time: A larger search region takes longer to traverse
but yields better results, since more similar patches can be found. We use same 
search region in both stages. The
temporal search window is fixed at $w_t = 5$. The spatial search region is set at
$w_x = 37$ for color videos and $w_x = 45$ for grayscale videos.
With these values, the search region captures the $5$ frame trajectory of the
target patch as long as it moves slower than $37/5 = 7.4$ pixels/frame (for
color videos).

@@

For color videos, we will set the spatial patch size at $s_x = 5$ for both
stages, whereas for grayscale videos we use $s_x = 7$. Larger patches (up to $s_x = 9$) lead to an mild improvement of the
results (on average). This improvement comes at a considerable increase in the
computation time, since our current implementation is $\mathcal O(s_x^4)$.
Thus, have found that a good compromise between computation time and quality is
given by $s_x = 5$.

@@

Due the coupling of the first and second stages, determining the optimal values
for the ranks $r_1, r_2$ and the number of patches $n_1,n_2$ requires a four
dimensional parameter search.
%
Instead, we fix the first stage parameters, both for color and grayscale videos
at
\[r_1 = s_{x,1}^2s_{t,1}\quad\quad n_{1} = 400,\]
which we found to give good results. In doing so, we are being conservative and
use the full rank covariance matrix to eliminate the least signal components.
%
We then run a parameter search to determine $r_2$ and $n_{2}$.
The results is the ones shown in Figure \ref{fig:rank1-nsim1-fpsnr}. From here
establish the following values as a function of $\sigma$:
%\[r_2(\sigma) = 40 - \frac13(\sigma - 10)\quad\quad n_{2} = 150.\]
\[r_2(\sigma) = 35 - \frac12(\sigma - 10)\quad\quad n_{2} = 100.\]
The same values are used for color and grayscale videos.
Note that a higher value for $n_{2}$ would give slightly worse results,
at a lower computational cost.

\begin{table}
	\centering
	\begin{tabular}{l | c c | c c }
		& \multicolumn{2}{c|}{color} 
		& \multicolumn{2}{c}{grayscale}\\
		Stage                            & 1st    & 2nd   & 1st   & 2nd   \\\hline\hline
		$s_x$        &  5 & 5 & 7 & 7 \\
		$s_t$        &  1--4  & 1--4  & 1--4  & 1--4  \\
		$w_x$        &  37    & 37    & 45    & 45    \\
		$w_x$        &   5    & 5     & 5     & 5    \\
		$n$          & 400    & 100   & 400   & 100  \\
		$r$          & full   & $35 - \frac12(\sigma - 10)$    & full  & $35 - \frac12(\sigma - 10)$   \\
%		Distance threshold $\tau$   & n/a  & 432  & n/a  & 432  & n/a  & 432 \\
%		Beta                        & 1    & 1.2  & 1    & 1.2  & 1    & 1.2 \\\hline
	\end{tabular}
	\caption{Selected parameters.}
	\label{tab:parameters}
\end{table}

@@

		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-curves_s10_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-curves_s20_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-curves_s30_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr_r1-np1-curves_s40_average_mono.eps}\\
	\end{center}
	\caption{Basic estimate PSNR, averaged over four sequences, when varying the
		second stage parameters $r_1$ and $n_1$. From left to right, results for
		$\sigma = 10, 20, 30, 40$. In the top row: results for RGB sequences, bottom
		row, results for grayscale sequences.} 
	\label{fig:rank1-nsim1-bpsnr}
\end{figure*}

@@

		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr-fpsnr_r1-np1-curves_s10_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr-fpsnr_r1-np1-curves_s20_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr-fpsnr_r1-np1-curves_s30_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/bpsnr-fpsnr_r1-np1-curves_s40_average_mono.eps}\\
	\end{center}
	\caption{PSNRs of final vs. basic estimate. Averages over four sequences, when varying the
		second stage parameters $r_1$ and $n_1$. From left to right, results for
		$\sigma = 10, 20, 30, 40$. Top: results for RGB sequences; bottom:
		results for grayscale sequences. The width of the line is proportional to
		the rank. \pcomment{These plots contradict the argument in favour of using
		full rank in the first stage (supposedly introducing a higher variance, but a
		smaller bias). It turns out, that (1) the final PSNR is pretty much
		correlated with the basic PSNR, and (2) the optimal rank for the final
		estimate is lower than that of the basic estimate. I do not have an explanation for this.}} 
	\label{fig:rank1-nsim1-bpsnr-fpsnr}
\end{figure*}

@@

		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r1-np1-curves_s10_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r1-np1-curves_s20_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r1-np1-curves_s30_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r1-np1-curves_s40_average_mono.eps}\\
	\end{center}
	\caption{Final estimate PSNR, averaged over four sequences, when varying the
		second stage parameters $r_1$ and $n_1$. From left to right, results for
		$\sigma = 10, 20, 30, 40$. In the top row: results for RGB sequences, bottom
		row, results for grayscale sequences.} 
	\label{fig:rank1-nsim1-fpsnr}
\end{figure*}

@@

		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-curves_s10_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-curves_s20_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-curves_s30_average_mono.eps}%
		\includegraphics[width=.25\textwidth]{figs/params_tables/fpsnr_r2-np2-curves_s40_average_mono.eps}\\

@@

@@

	\caption{PSNRs obtained for the classic color test sequences. See text for
		details. \pcomment{NOTE: the values in the table need to be updated!!}}

@@

	\caption{Quantitative denoising results for some classic grayscale test sequences. See text for details. \pcomment{NOTE: the values in the table need to be updated!!}}

@@

In this section we present results obtained with the proposed
method on some classical test sequences with Gaussian noise added.
We consider separatedly the cases of grayscale and color videos.
All results were computed with the parameters shown in Table \ref{tab:parameters}.

@@

	Sintel's database. See text for details.
	\pcomment{NOTE: the values in the table need to be updated!!}}
